Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
editing 1 pt mid edits

In this subsection, we restrict attention to the case of effective divisors.

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \br^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$ (i.e. the taxicab metric).
Let $\mss_i := \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\sum_{i = 0} ^{m + 1} v_i = c_i\}$ be the set of points in
$\bz_{\geq 0}^{m + 1}$ with sum of components equal to $c_i$.
\end{defn}

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \br^n$ be two vectors. Let $i$
be the right-most coordinate such that $v_i$ is nonzero
and $j$ be the left-most coordinate such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

\begin{thm}
\label{thm:proj-one-point}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \alpha H$, with $\alpha = \frac{p}{q} \in \bq$
and $H = V(x_k)$ a hyperplane of $\bp^m$, be a divisor on $\bp^m$.
Let
\[
	0 = \frac{c_0}{d_0} <
	\frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{p}{q}
\]

\noindent
be the best lower approximations of $\alpha$. Then the
canonical ring
\[
	R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)
\]

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $\vec{v} \in \bz_{\geq 0}^{m + 1}$
such that $\sum_{j = 0}^{m} v_j = c_i$. Furthermore, it has
relations of the following two types.

For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\substack{\vec{y} \in \mss_{h_{i, j}} \\ \vec{z} \in
	\mss_{h_{i, j} + 1}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}} \;
	(F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*}

Furthermore, there are the relations for each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

>>>>

In this subsection, we restrict attention to the case of effective fractional divisors $D \in \bq \otimes \di \bp^m$.

\begin{defn}
\label{defn:vec-sum}
For $\vec{v} \in \br^n$, denote $\deg \vec{v} := \sum_{i = 0}
^n v_i$.
Let 
\begin{align*}
	\mss_i := \left \{\vec{v} \in \bz_{\geq 0}^{m + 1} \; : \;
\deg \vec v = c_i \right\}.	
\end{align*}
\end{defn}

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \br^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

\begin{thm}
\label{thm:proj-one-point}
Let $\bp^m \cong \proj \bk [x_0, \ldots, x_m].$ Let $D = \alpha H$, with $\alpha = \frac{p}{q} \in \bq$
and $H := V(x_k)$ a hyperplane of $\bp^m$.
Let
\[
	0 = \frac{c_0}{d_0} <
	\frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{p}{q}
\]

\noindent
be the best lower approximations of $\alpha$. Then, the
canonical ring
\[
	R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)
\]

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $1 \leq i \leq r$, $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, it has
relations of the following two types:
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\substack{\vec{y} \in \mss_{h_{i, j}} \\ \vec{z} \in
	\mss_{h_{i, j} + 1}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}} \;
	(F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*}
	\item For each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) there is a relation of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

\end{enumerate}
